---
title: "Impact of Intervention on Conflict Outcomes"
author: "K. Heger and B. Abendroth"
date: "13 November 2015"
output: 
  pdf_document: 
    latex_engine: xelatex
bibliography: BibliographyA3.bib
---
# Introduction
In policy-making, military intervention is often depicted as a solution to internal conflicts, while at the same time, academic research has revealed that intervention oftentimes rather prolongs and exacerbates civil wars. Our research project looks at how different types of third-party intervention (military, non-military, troops) into intrastate wars affect the outcome of the conflict (rebel or government victory, negotiated settlement, low intensity conflict). This paper briefly outlines the data gathering and cleaning process, before conducting basic descriptive statistics and a first regression analysis. A more comprehensive analysis will follow in the final paper. 

# Data Gathering
We use as a baseline for our analysis a publicly available dataset created by Sullivan and Karreth (@sullivan2014conditional), which  will henceforth be refered to as the CIMI Dataset. Sullivan and Karreth draw from the 3.3 version of the Non-state Actors (NSA) dataset (@cunningham2013non), as well as the University of Uppsala Conflict Data Program's (UCDP) Conflict Termination Dataset (@kreutz2010and) and External Support Dataset (@hogbladh2011external).

## Dependent Variable
We adopt our dependent variable conflict_outcome from the CIMI Dataset, which disaggregates four different conflict outcomes coded from the UCDP Conflict Termination Dataset. Internal conflicts either fade out into a period of low activity (0), they end with a victory by the rebels (1), they terminate with some form of negotiated settlement (2), or with government victory (3). 

## Independent Variables
Measuring the impact of third-party military intervention on conflict outcomes, we distinguish between troops, military, and non-military support, as coded by the NSA Dataset. Further key explanatory variables in determining conflict outcomes are the capacity of rebels and government relative to each other. CIMI includes both rebel military capacity relative to the government and the log of GDP per capita as a measure for government strength. We borrow the Cold War dummy, the coup dummy as well as the log of conflict duration as controls from the CIMI dataset. 

# Data Cleaning And Merging
Considering that we drew variables from two different datasets (CIMI and NSA), we had to clean and merge these two datasets. 
We first created a subset of the CIMI Dataset with the relevant variables needed for our statistical model, which we called Cleaned_CIMI. As is often the case with conflict data, we were faced with a missing values problem and deleted all observations with missing values in the dependent variable, leaving us with a data frame with 425 observations and 10 variables.
In a second step, we cleaned the NSA Dataset for our purpose and equally created a subset of the data frame called Cleaned_NSA. We coded two dummies, external support du rebels and external support to governments, from the existing character variables. In addition, we transformed the character variables describing the type of support received by governments and rebels into categorical variables with the following factors: 1=troops, 2=military, 3=non-military. Missing data on these covariates reduced the number of observations from 578 to 160. 
Lastly, we merged the cleaned datasets around the unique identifier dyadid. As is common in recent conflict research, both CIMI and NSA datasets take conflict dyads as their unit of analysis, modeling civil wars as dyadic struggles between the government and an opposition party, which enables easy merging. The merged dataset Dataset_1 has 497 observations and 23 variables. 

# Data Analysis
In the following, we conduct some basic descriptive as well as inferential statistics. 

## Descriptive Statistics
Figure 1 shows the frequency of the different conflict outcomes. The most common outcome for a conflict dyad is low activity, defined as a violent conflict with less than 25 deaths per year for at least one year. While 44% of all conflict dyads level off like that, 35% end in a victory of one side. Governments win the conflict in 22% of the cases, rebels prevail in only 13%. 21% of conflicts terminate with a negotiated settlements.  
Figure 2 displays the frequency, with which different types of support are lent to the government. In 20% of the cases, governments are supported with troops, in 72% they receive some other form of military support. Only in 7% of the cases, third-party governments support the government with non-military means.   
Figure 3 depicts the kind of support rebels receive. Similar to support to governments, by far the most common type of support is military support with 70%. Both troops and non-military support are less frequent with 14% and 15%, respectively. The variable recording the type of support to rebels further includes as a fourth category endorsement. In 2 cases in our sample, foreign governments supported the rebels by publicly endorsing them. 

0: Low Activity  
1: Rebel Victory  
2: Settlement  
3: Government Victory  

Figure 2 and 3 show the frequency of the different types of support to a government or a rebel group respectively. Both refer to the following types of support:

1: Troops  
2: Military  
3: Non-military  


```{r fig.width=5, fig.height=4, echo=FALSE}
library("ggplot2")
Dataset_1 <- read.csv("https://raw.githubusercontent.com/KatrinHeger/CollaborativeResearchProject/master/Assignment3_Data-Analysis/Dataset_1_outcome_HG_RMD.csv")

qplot(Dataset_1$conflict_outcome, geom = "histogram", binwidth = .5, main = "Fig.1: Frequency of Conflict Outcomes", xlab = "Outcomes", ylab = "Frequency", fill = I("lightblue"))
```

```{r fig.width=5, fig.height=4, echo=FALSE}
Dataset_1_gtype_HG <- read.csv("https://raw.githubusercontent.com/KatrinHeger/CollaborativeResearchProject/master/Assignment3_Data-Analysis/Dataset_1_gtype_HG_Rmd.csv")

qplot(Dataset_1_gtype_HG$gtypesup_cat, geom = "histogram", binwidth = .5, main = "Fig.2: Frequency of Government Support Types", xlab = "Support Types", ylab = "Frequency", fill = I("lightgreen"))
```


```{r fig.width=5, fig.height=4, echo=FALSE}
Dataset_1_rtype_HG <- read.csv("https://raw.githubusercontent.com/KatrinHeger/CollaborativeResearchProject/master/Assignment3_Data-Analysis/Dataset_1_rtype_HG_Rmd.csv")

qplot(Dataset_1_rtype_HG$rtypesup_cat, geom = "histogram", binwidth = .5, main = "Fig.3: Frequency of Rebel Support Types", xlab = "Support Types", ylab = "Frequency", alpha = I(.9), fill = I("brown"))
```


## Multinomial Logistical Regression
Our baseline model estimates a multinomial logit model because our dependent variable - conflict outcome - is a categorical variable with the four outcomes government victory, rebel victory, negotiated settlement, and low intensity conflict. The results tell us how different kinds of third-party intervention affect the likelihood of a conflict having a specific outcome vs. the reference (low intensity conflict).

\begin{equation}\begin{aligned} Pr(Y_{i}=j) = \frac{exp(X_{i}\beta_{j})} {\sum_{j=1}^{J} exp(X_{i}\beta_{j})}\end{aligned}\end{equation}

In our model, j represents each of the four outcomes and i indixes the unit of analysis, being a dyadic episode of intrastate armed conflict. FOllowing Sullivan and Karreth, we omit the most common conflict outcome, low intensity conflict, as the baseline category and calculate estimates for the remaining outcomes government victory, negotiated settlement and rebel victory.


The following Tables 1-5 show the results for our five different logit models.

```{r echo=FALSE, results='asis', warning=FALSE, message=FALSE}
library("nnet")
library("memisc")
library("stargazer")

Dataset_1_Multinom <- read.csv("https://raw.githubusercontent.com/KatrinHeger/CollaborativeResearchProject/master/Assignment3_Data-Analysis/Dataset_1_Multinom.csv")

dat.1 <- Dataset_1_Multinom[is.na(Dataset_1_Multinom$conflict_outcome) == FALSE & is.na(Dataset_1_Multinom$rebel.support_d) == FALSE, c("conflict_outcome", "rebel.support_d")]
m1 <- multinom(conflict_outcome ~ rebel.support_d, data = dat.1)

stargazer::stargazer(m1, digits = 2)
```


```{r, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
library("nnet")
library("memisc")
library("stargazer")

Dataset_1_Multinom <- read.csv("https://raw.githubusercontent.com/KatrinHeger/CollaborativeResearchProject/master/Assignment3_Data-Analysis/Dataset_1_Multinom.csv")

dat.2 <- Dataset_1_Multinom[is.na(Dataset_1_Multinom$conflict_outcome) == FALSE & is.na(Dataset_1_Multinom$rebel.support_d) == FALSE & is.na(Dataset_1_Multinom$gov.support_d) == FALSE, ]
m2 <- multinom(conflict_outcome ~ rebel.support_d + gov.support_d, data = dat.2)

stargazer::stargazer(m2, digits = 2)
```


```{r, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
library("nnet")
library("memisc")
library("stargazer")

Dataset_1_Multinom <- read.csv("https://raw.githubusercontent.com/KatrinHeger/CollaborativeResearchProject/master/Assignment3_Data-Analysis/Dataset_1_Multinom.csv")

dat.3 <- Dataset_1_Multinom[is.na(Dataset_1_Multinom$conflict_outcome) == FALSE & is.na(Dataset_1_Multinom$rebel.support_d) == FALSE & is.na(Dataset_1_Multinom$gov.support_d) == FALSE & is.na(Dataset_1_Multinom$rtypesup_cat) == FALSE & is.na(Dataset_1_Multinom$gtypesup_cat) == FALSE, ]
m3 <- multinom(conflict_outcome ~ rebel.support_d + gov.support_d + rtypesup_cat, data = dat.3)

stargazer::stargazer(m3, digits = 2)
```


```{r, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
library("nnet")
library("memisc")
library("stargazer")

Dataset_1_Multinom <- read.csv("https://raw.githubusercontent.com/KatrinHeger/CollaborativeResearchProject/master/Assignment3_Data-Analysis/Dataset_1_Multinom.csv")

dat.4 <- Dataset_1_Multinom[is.na(Dataset_1_Multinom$conflict_outcome) == FALSE & is.na(Dataset_1_Multinom$rebel.support_d) == FALSE & is.na(Dataset_1_Multinom$gov.support_d) == FALSE & is.na(Dataset_1_Multinom$rtypesup_cat) == FALSE & is.na(Dataset_1_Multinom$gtypesup_cat) == FALSE, ]
m4 <- multinom(conflict_outcome ~ rebel.support_d + gov.support_d + rtypesup_cat + gtypesup_cat, data = dat.4)

stargazer::stargazer(m4, digits = 2)
```


```{r, echo=FALSE, results='asis', warning=FALSE, message=FALSE}
library("nnet")
library("memisc")
library("stargazer")

Dataset_1_Multinom <- read.csv("https://raw.githubusercontent.com/KatrinHeger/CollaborativeResearchProject/master/Assignment3_Data-Analysis/Dataset_1_Multinom.csv")

dat.5 <- Dataset_1_Multinom[is.na(Dataset_1_Multinom$conflict_outcome) == FALSE & is.na(Dataset_1_Multinom$rebel.support_d) == FALSE & is.na(Dataset_1_Multinom$gov.support_d) == FALSE & is.na(Dataset_1_Multinom$rtypesup_cat) == FALSE & is.na(Dataset_1_Multinom$gtypesup_cat) == FALSE & is.na(Dataset_1_Multinom$fightcaphigh) == FALSE & is.na(Dataset_1_Multinom$lngdp) == FALSE & is.na(Dataset_1_Multinom$lnyears) == FALSE & is.na(Dataset_1_Multinom$postCW) == FALSE, ]
m5 <- multinom(conflict_outcome ~ rebel.support_d + gov.support_d + rtypesup_cat + gtypesup_cat + fightcaphigh + lngdp + lnyears + postCW, data = dat.5)

stargazer::stargazer(m5, digits = 2)
```
## Outlook
Our initial analysis has shown that neither support for the rebels nor for the government is significant. Other tables showed signigicance for some variables. Table 3, for example, indicates that rebel endorsement by a third party is significant over all models, as well as non-military support for the rebels. However, due to the structure of the multinomial logit estimation, it is very complicated to interprete the effects from the above inserted tables. In future analysis, we thus will have to evaluate the statistical significance of each independent variable. Moreover, intervention into a conflict and military strength might vary within one conflict dyad. This variation could be accounted for by switching the unit of analysis from conflict dyads to conflict dyad years. 

## R Packages
In our analysis, the following packages have been used: @CiteR, @rio, @warnes2014gdata, @croissant2012estimation, @wickham2014dplyr, @fox2009car, @wickham2009ggplot2, @hlavac2013stargazer, @ripley2015package, @testdat_package, @harrell2006hmisc, @elff2015package.

## Bibliography
